---
title: "QuickStart"
author: "Markus List and Konstantinos Mechteridis"
output: html_document
---



## QuickStart KeyPathwayMiner R-package {.tabset}
This is an R Markdwon document on how to get started with the KeyPathwayMiner R-package.
We will go through how to execute local and remote calls of KeyPathwayMiner and how to
change and view the run parameters.
For more details on using the KeyPathwayMiner please visit our website: <https://exbio.wzw.tum.de/keypathwayminer/>.

### Prepare data
```{r prepare_data}
source(system.file("inst/extdata/api_testing/", "matrix-hd-up.dat", package = "Rkpm"))
# KeyPathWayMiner URL
url <- "https://exbio.wzw.tum.de/keypathwayminer/"

# Load indicator matrices as data.frame
huntington_disease_up <- as.data.frame.matrix(read.delim("matrix-hd-up.dat", header = FALSE))

huntington_disease_down <- as.data.frame.matrix(read.delim("matrix-hd-down.dat", header = FALSE))

huntington_list <- list(huntington_disease_up, huntington_disease_down)

# Get overview of available networks
print(get_networks(url))

# Get ID of I2D Homo_sapiens entrez network
I2D.id <- get_networks(url)[["I2D Homo_sapiens entrez"]]

```
### Remote runs
### Local runs



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
