---
title: "KeyPathwayMineR - R Package for the web and standalone version of KeyPathwayMiner"
date: FALSE
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{KeyPathwayMineR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---
```{r, include = FALSE}
# General markdown options
# Collapses output and comments it to
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r quick_load_data, include = FALSE}
#library(KeyPathwayMineR)

# Initialize for testing
# Path must be to the working direcotry where the package is located
devtools::load_all(path = "~/Workspace/keypathwayminer/keypathwayminer-R/")
.jinit("../inst/java/keypathwayminer-standalone-5.0.jar")

# Read huntington_disease_up as data.frame
huntington_disease_up <- as.data.frame.matrix(read.delim(
  system.file(
    package = "KeyPathwayMineR",
    "extdata/datasets",
    "huntington-gene-expression-UP.txt"
  ),
  header = TRUE
))
  
# The entries of the huntington_disease_up dataset are p-values. To create an indicator matrix the to_indicator_matrix()
# functio can be used
huntington_disease_up <- to_indicator_matrix(
  numerical_matrix = huntington_disease_up,
  operator = "<", threshold = 0.005
)

huntington_disease_down <- as.data.frame.matrix(read.delim(
  system.file(
    package = "KeyPathwayMineR",
    "extdata/datasets",
    "huntington-gene-expression-DOWN.txt"
  ),
  header = FALSE
))

sample_network <- system.file("extdata", "sampleNetwork.sif", package = "KeyPathwayMineR")
```
# Introduction
## KeyPathwayMiner
Given a **biological network** and a **set of case-control studies**, KeyPathwayMiner(KPM) efficiently extracts all             **maximal connected sub-networks**. These sub-networks contain the genes that are **mainly** dysregulated, e.g.,           differentially expressed, in **most** cases studied:

   * The exact quantities for “mainly” and “most” are modeled with two easy-to-interpret parameters **(K, L)** that allows      the user to control the number of outliers (not dysregulated genes/cases) in the solutions.
     
Two different approaches for extracting subnetworks that are enriched for active/deregulated genes have been               implemented:

  * **INES:**  Extract all maximal sub-networks containing nodes with no more than L inactive cases (0's) besides of K                    exceptions.
      
  * **GLONE:** Extracts maximal sub-networks where the total sum of **not**-active/diff. exp. cases is at most L. 
      
For more information please visit the [KPM website](https://exbio.wzw.tum.de/keypathwayminer/).

## KeyPathwayMineR
The following package provides an easy way for using KPM and visualize the extracted subnetworks via R.
The user can decide whether he wants to call KPM on his computer (via standalone) or with the Web API. 
With the `kpm_options()` object the arguments for the execution of the program can be easily processed and viewed.
Follow the instructions on [KeyPathwayMineR github](https://github.com/baumbachlab/keypathwayminer-R) to install the package.

# Data preparation
In the following section, we will look at how to prepare data for a KeyPathwayMiner execution.
Subsequently, we will present the options manager `kpm_options()`, which allows the user to set the run parameters.  

## Input data
Two types of  input files are necessary to run KeyPathwayMineR:

 * **matrix file/s** from the expression studies. Either numeric (e.g. p-values or fold changes) or as indicator matrix.
 * **graph file** containing all the interactions of the protein interaction network. 
 
The data can be provided as data.frames or as files. For more information on the correct format of the input files, use `vignette("input_files_format")`.

### Dataset file/s
Lets get started with the dataset files (matrix files):
```{r datasets_files_a}
huntington_disease_down <- as.data.frame.matrix(read.delim(
  system.file(
    package = "KeyPathwayMineR",
    "extdata/datasets",
    "huntington-gene-expression-DOWN.txt"
  ),
  header = FALSE
))
```
The entries of the huntington_disease_up dataset are p-values. To create an indicator matrix the to_indicator_matrix() function can be used.
```{r datasets_files_b}
# Read huntington_disease_up as data.frame
huntington_disease_up <- as.data.frame.matrix(read.delim(
  system.file(
    package = "KeyPathwayMineR",
    "extdata/datasets",
    "huntington-gene-expression-UP.txt"
  ),
  header = TRUE
))

# Convert to indicator matrix
huntington_disease_up <- to_indicator_matrix(
  numerical_matrix = huntington_disease_up,
  operator = "<", threshold = 0.005
)
```
You could also provide the path to an indicator matrix file. Assuming the file has the correct format KeyPathwayMineR will
automatically detect the type of input and processes it according to the execution type (local or remote) (see `vignette("input_files_format")`).

### Graph file
In the next step, we will go through how to prepare the network files for the execution.
Firstly the graph_file can be provided as a path:
```{r graph_file_path}
sample_network <- system.file("extdata",
  "sampleNetwork.sif",
  package = "KeyPathwayMineR"
)
```
For the remote execution, the user can as well select a graph_id from the web service.
```{r get_networks}
# Get overview of available networks
get_networks()
```
Now the user just needs to save the graph_id from the graph he wants to use:
```{r graph_id}
kpm_options(graph_id = 10)
```


## Options
The options manager provides the user with a variety of arguments that can be customized for the execution. To get a documentation and an overview of all parameters available, you can run the following commands
```{r options_manager, eval = FALSE}
# For the documentation
?kpm_options()

# For a  general overview of all parameters use without head
kpm_options()
```
You can change the options by choosing the option you want and assigning it to a new value as follows:
```{r change_option}
# Initial value
kpm_options()$algorithm

# Change algorithm to ACO
kpm_options(algorithm = "ACO")

# After changing the value
kpm_options()$algorithm

# You can also change multiple parameters by seperating them with a comma
kpm_options(algorithm = "Optimal", execution = "Local")
```
Reset your options for a new run like that:
```{r reset_option}
# Initial value
reset_options()
```

# Execution and visualization
Once all the input files are prepared, we are ready to start the pathway enrichment analysis. This section will demonstrate how to run the local and remote execution of KeyPathwayMiner and how to visualize the acquired results with shiny.

## Local
For users who want to conduct extensive analysis, the local execution is the way to go. The user can specify different case exception parameters (L) for every dataset. Furthermore, he can also analyze the robustness of the provided network by choosing one of the four perturbation techniques.

For the users who want to go even more into detail, we provide a properties file where the user can specify the default parameters for the local run. To access and edit the properties file use `edit(file = system.file(package = "KeyPathwayMineR", "inst/extdata/kpm.properties"))`. Keep in mind that options set with kpm_options() have higher priority than the options in the kpm.properites file.

In the first example, we will perform an unranged run using INES. In addition to that we will analyse the robustness of the network with the perturbation technique "nodeswap" and the following parameters:
```{r local_example_1, eval = FALSE}
# Use perturbation
kpm_options(
  execution = "Local", strategy = "INES", algorithm = "Greedy",
  l_min = 20,
  k_min = 5,
  perturbation_technique = "nodeswap",
  with_perturbation = TRUE,
  remove_bens = TRUE,
  perturbation_start = 10,
  perturbation_step = 10,
  perturbation_max = 20,
  graphs_per_step = 2
)

# Execute remote run by using a custom graph_file
local_example_1 <- kpm(graph = sample_network, indicator_matrices = huntington_disease_up)

# Visualize the results with shiny
visualize_result(local_example_1)
```
In the next example, an unranged GLONE run will be demonstrated for multiple datasets. Since we are using two datasets we will have to define two l parameters. This is possible by providing a vector as input for l_min:
```{r local_example_2, eval = FALSE}
# Reset settings
settings::reset(kpm_options)

indicator_matrices <- huntington_disease_up 
# Combine multiple datasets:
kpm_options(
  execution = "Local",
  strategy = "GLONE",
  algorithm = "Greedy",
  l_min = 15,
  l_step = 5,
  l_max = 50,
  k_min = 2,
  use_range_l = TRUE,
  user_range_k = TRUE
)

# Run kpm
local_example_2 <- kpm(graph = sample_network, indicator_matrices = indicator_matrices)

# Visualize the results with shiny
visualize_result(local_example_2)
```
In the last example, we will perform a ranged (batch) run using INES. By setting `use_range_l = TRUE` you 
specify that the parameter L.

**Example:** `l_min = 4`, `l_max = 8` and `l_step = 2` would mean that the KPM will run with L = 4, L = 6 and L = 8.

```{r local_example_3, eval = FALSE}
settings::reset(kpm_options)

indicator_matrices <- list(huntington_disease_up, huntington_disease_down)

# Use ranged values with batch
kpm_options(
  execution = "Local", strategy = "INES", algorithm = "Greedy",
  use_range_l = TRUE,
  l_min = c(20, 15),
  l_step = c(2, 1),
  l_max = c(24, 17),
  k_min = 5,
  link_type = "AND"
)

# Run kpm
local_example_3 <- kpm(graph = sample_network, indicator_matrices = indicator_matrices)

# Visualize the results with shiny
visualize_result(local_example_3)
```
When `use_range` is set to **FALSE** (by default) only `l_min` is relevant. The same applies to K. 

## Remote
The remote execution is convenient when the user does not have sufficient resources to perform an analysis and wants to perform a simple analysis.

A single dataset with fixed parameters for K and L, and the INES Algorithm was selected in the first example. `async = FALSE`  means that we have a blocking request. Blocking, suggesting that the user will have to wait until the execution is complete. For this run, we will use the graph_id 10, which is the "I2D Homo_sapiens entrez" graph.
```{r remote_example_1, eval = FALSE}
settings::reset(kpm_options)
# Configure options for the run
kpm_options(
  execution = "Remote",
  async = FALSE,
  strategy = "INES",
  remove_bens = TRUE,
  algorithm = "Greedy",
  l_min = 20,
  k_min = 5
)

# Start run with huntington_disease_up dataset
remote_example_1 <- kpm(indicator_matrices = huntington_disease_up, 
                        graph = sample_network)

# Visualize the results with shiny
visualize_result(remote_example_1)

# Open the result page where you can  monitor the progress of both tasks
browseURL(get_result_url(remote_example_1))
```
In the next example for the remote execution, we will take a look at how to combine and use multiple datasets. 
We will start a run with INES, K = 1 and L = 10% for the two datasets huntington_disease_up and huntington_disease_down.
You may have noticed that the link_type was set to OR. This means the two datasets will be logically connected with the "OR" operator.
```{r remote_example_2, eval = FALSE}
# Reset settings
settings::reset(kpm_options)

kpm_options(
  execution = "Remote",
  async = FALSE,
  strategy = "INES",
  algorithm = "Greedy",
  remove_bens = TRUE,
  l_same_percentage = TRUE, same_percentage = 10,
  k_min = 1,
  graph_id = 10,
  link_type = "OR"
)

# Add the two datasets to a list
indicator_matrices <- list(huntington_disease_down, huntington_disease_up)
# Run kpm
remote_example_2 <- kpm(indicator_matrices = indicator_matrices)

# Visualize the results with result_remote_example_3
visualize_result(remote_example_2)

# Open the result page where you can  monitor the progress of both tasks
browseURL(remote_example_2)
```  
**Important note:** The **web** service does not allow individual fixed (L case-exceptions) parameters to be set for each matrix at the moment. Robustness analysis is also not supported via Remote execution. If you want to use batch runs and robustness analysis, please use the **Local** execution.

# Conclusion
Our goal is to simplify the user's work process when using KeyPathwayMiner. Users can perform their analysis directly in R without having to worry about how to access Jar files or set up HTTP requests. 
At the same time, R provides an efficient way to prepare, handle and further process data. This is for example very practical for the preparation of indicator matrices and the visualization of the extracted results. Finally, the user can choose whether he wants to use the remote or the local version depending on the performance of his computer and his requirements.
